\chapter{Solovay Model}

The Solovay model is an interesting model in which choice fails.
It is of particular intrigue to analysts due to its many interesting analytical and measure-theoretic properties.
The main results were established in the Spring of 1964, and were presented at the July meeting of the Association of Symbolic Logic at Bristol, England \autocite[see footnote * in][1]{solovay1970}.
The methods and arguments were developed just a year after Cohen's creation of forcing,
and were highly sophisticated for its time.

On top of introducing the Solovay model, \autocite{solovay1970} also introduced the method of generic filters and Boolean valued models,
which was a major refinement towards Cohen's original techniques of forcing.
In contrast to Cohen's syntactic approach, known as ramified forcing, Solovay's methods were noticeably more algebraic and abstract.

\section{Lévy Collapse}

\newcommand*{\Lv}{\operatorname{Lv}}
\begin{definition}
    For any \(S\subseteq \mathbf{ORD}\), the \emph{Lévy collapsing order} for \(S\) is
    \[
        \Lv(S) = \set{
            \begin{split}
                p: &\ p\text{ is a function} \\
                   &\land \abs{p} < \omega \\
                   &\land \dom(p) \subseteq S\times\omega \\
                   &\land \forall \Angle{\alpha,n}\in\dom(p)\sq{p(\alpha,n) = 0 \lor p(\alpha,n)\in \alpha}
            \end{split}
        }
    \]
    ordered by reverse inclusion.
\end{definition}
\begin{remark}
    Forcing with the Lévy collapse will add surjections \(\omega\to\alpha\) for each \(\alpha\in S\).
    The canonical application of this is forcing with \(\Lv(\kappa)\), where \(\kappa\) regular,
    then it will follow that \(M[G]\models \kappa = \omega_1\).
\end{remark}

We have an immediate consequence of the definition.
% move this?
\begin{proposition}
    \(\Lv(S)\) is almost homogeneous.
\end{proposition}
\begin{proof}
    Let \(p,q\in\Lv(S)\), we can find a bijection \(f:\omega\to\omega\) such that
    for all \(\Angle{\alpha,m}\in\dom(p), \Angle{\beta,n}\in\dom(q)\), we have \(f(m) \ne n\).
    We can extend \(f\) into an automorphism \(i:\P\to\P\) as follows
    \[ i(r) = \paren{\alpha,m}\mapsto r\paren{\alpha, f^{-1}(m)} \]
    where \(i(r)\paren{\alpha,m}\) is undefined when \(r\paren{\alpha,f^{-1}(m)}\) is.
    Then by our selection of \(f\), we can see that \(\dom(i(p)) \cap\dom(q) = 0\) so \(i(p)\) and \(q\) are compatible.
\end{proof}

\begin{lemma}[10.17 in {\autocite[127]{kanamori2003}}] \label{lemma:levy_elementary}
    Elementary properties of the Lévy collapsing order,
    \begin{enumerate}[(a)]
        %\item \(\Lv(S)\) is \(\lambda\)-closed.
        \item Suppose that \(S = X\cup Y\) is a disjoint union,
            and set \(\P_0 = \Lv(X)\) and \(\P_1 = \Lv(Y)\).
            Then \(G\) is \(\Lv(S)\)-generic iff \(G\) is of the form
            \[G = \set{p\cup q: p\in G_0\land q\in G_1 }\]
            where \(G_0\) is \(\P_0\)-generic, \(\paren{\P_1}^{M[G_0]} = \P_1\),
            and \(G_1\) is \(\P_1\)-generic over \(M[G_0]\).
        \item If \(\kappa\) is uncountable and regular, %\(\kappa > \lambda\) and either \(\kappa\) is inaccessible or \(\lambda = \omega\),
            then \(\Lv(\kappa)\) has the \(\kappa\)-c.c.
        \item Forcing with \(\Lv(\kappa)\) when it has \(\kappa\)-c.c. will preserve cardinals \(\geq \kappa\).
        \item Suppose \(\kappa\) is regular, \(\Lv(\kappa)\) has the \(\kappa\)-c.c., and
            \(G\) is \(\Lv(\kappa)\)-generic.
            Then for any \(s\in M[G]\) with \(s:\gamma \to \mathbf{ORD}\) where \(\gamma < \kappa\),
            there is a \(\delta<\kappa\) such that \(s\in M\sq{G\cap\Lv(\delta)}\).
    \end{enumerate}
\end{lemma}

\begin{remark}
%    Part (a) is result of Solovay's demonstration that forcing with a product poset is basically iterated forcing.

    Part (d) tells us that when forcing with \(\Lv(\kappa)\),
    the collapse can be decomposed into stages, and
    any \(<\kappa\) sequence of ordinals will be introduced at some earlier stage.
    This will be made precise in Solovay's ``important lemma''.
\end{remark}


Another ingredient of Solovay's important lemma is this characterisation of \(\Lv(\set{\alpha})\),
this characterisation can be thought of as a partial converse to the collapsing nature of \(\Lv(\set{\alpha})\),
it says that a partial order which collapses \(\alpha\) into \(\omega\) is ``basically'' \(\Lv(\set{\alpha})\)
in the sense that both posets structurally share a common dense subset.
More precisely,
\begin{lemma}\label{lemma:characterisation_collapse}
    Suppose \(M\models \abs{\P} \leq \abs{\alpha}\) and
    \[ 1_\P \forces \exists f\paren{f:\omega\to\alpha \text{ is onto}}.\]
    Then there is a dense subset \(D_\alpha\subseteq \Lv(\set{\alpha})\)
    and an injective embedding \(D_\alpha \to \P\) whose image is dense.
\end{lemma}
\begin{proof}
    Set \(\nu = \abs{\alpha}\),
    note that combinatorially, we have for each \(p\in\P\) there is a maximal antichain below \(p\) of cardinality \(\nu\).

    The dense subset of \(\Lv(\set{\alpha})\) is
    \[ D_\alpha = \set{p: \exists n\in\omega \paren{p: \set{\alpha}\times n\to \alpha}}.\]

    Let \(\mathring{g}\) be name for the surjection \(g:\omega\to G\) that will be added, where \(G\) is \(\P\)-generic.
    Define an injective, dense embedding \(e: D_\alpha\to\P\) by recursion on \(\abs{p}\).
    Set \(e(0) = 1_\P\).
    Suppose \(e(p)\) has already been defined for some \(p\in D_\alpha\) with \(\dom(p) = \set{\alpha}\times n\),
    let \(\Angle{a_\xi^p: \xi<\alpha}\) enumerate a maximal antichain below \(e(p)\) of cardinality \(\nu\).
    Without loss of generality, it can be assumed that for each \(\xi\),
    \[ a_\xi^p \forces \mathring{g}(\check{n}) = \check{r} \text{ for some } r\in \P. \]
    Now set \(e(p\cup \set{\Angle{(\alpha, n), \xi}}) = a_\xi^p\) for each \(\xi<\alpha\).

    We can see that \(e\) is injective.
    To show that \(e''D_\alpha\) is dense, let \(r\in \P\).
    It is always the case that \(r\forces \check{r}\in\mathring{G}\),
    as we assume \(\P\) is atomless, we refine this condition and produce a \(s\leq r\)
    such that \(s\forces \mathring{g}(n) = \check{r}\) for some \(n\in\omega\).

    By construction of \(e\), there is a \(p\in D_\alpha\) with \(\abs{p} = n+1\) such that \(e(p)\) is compatible with \(s\).
    But also by construction, \(e(p)\) has already decided \(\mathring{g}(n)\) which means that \(e(p)\) must be below \(r\).
\end{proof}

\begin{remark}
    The conclusion of the lemma shows that \(\P\) and \(\Lv(\set{\alpha})\) are structurally similar,
    so if \(G\) is \(\P\)-generic, then there exists a \(\Lv(\set{\alpha})\)-generic filter \(H\) with \(M[G] = M[H]\),
    and vice-versa.
\end{remark}


\newcommand*{\Q}{{\mathbb{Q}}}
We now prove a key technical fact about the generic extension of the Lévy collapse
which allows us to enlarge the ground model of Lévy's construction to absorb any countable sequence of ordinals in the extension.
\begin{lemma}[Solovay's ``important lemma''] \label{lemma:important}
    Suppose \(\kappa>\omega\) is regular and \(G\) is \(\Lv(\kappa)\)-generic.
    For any \(s:\omega\to\mathbf{ORD}\), \(s\in M[G]\),
    there is a \(\Lv(\kappa)\)-generic filter \(H\) over \(M[s]\)
    such that \(M[G] = M[s][H]\).
\end{lemma}
\begin{proof}
    Apply \Cref{lemma:levy_elementary} (d), let \(\delta<\kappa\) such that \(s \in M[G\cap\Lv(\delta)]\).
    Decompose the collapse into stages by defining
    \begin{align*}
        G_0 &= G\cap\Lv(\delta), \\
        G_1 &= G\cap\Lv(\set{\delta}), \\
        G_2 &= G\cap\Lv(\kappa\setminus(\delta+1)).
    \end{align*}

    By \Cref{lemma:levy_elementary} (a), \(G_0\) is generic over \(\Lv(\delta)\),
    applying intermediate model property (\Cref{proposition:intermediate_model_property})
    we have \(M[G_0] = M[s][H_0]\) where \(\P\in M[s]\) and \(H_0\) is generic over \(\P\).

    Let \(\Q = \P\times\Lv(\set{\delta}) \in M[s]\),
    applying Product Lemma (\Cref{lemma:iterated_forcing_product}), \(M[s][H_0][G_1]\) is \(\Q\)-generic over \(M[s]\).
    The proof of Product Lemma actually gives an upper bound on the cardinality of \(\Q\), so
    \[ \abs{\Q} \leq \abs{\Lv(\delta)\times\Lv(\set{\delta})} = \abs{\Lv(\delta+1)} = \abs{\delta} \]
    and since \(\Q\) contains a copy of \(\Lv(\set{\delta})\),
    \[ 1_\Q \forces \exists f\paren{f:\omega\to\delta \text{ is onto}}. \]
    Applying \Cref{lemma:characterisation_collapse}
    there is a \(\Lv(\set{\delta})\)-generic (over \(M[s]\)) filter \(H_1\) such that \[ M[s][H_1] = M[s][H_0][G_1].\]

    As \(\Lv(\delta+1)\) also collapses \(\delta\), that is \( 1_{\Lv(\delta+1)} \forces \exists f\paren{f:\omega\to\delta \text{ is onto}} \),
    there is a \(\Lv(\delta+1)\)-generic (over \(M[s]\)) filter \(H_2\) such that \[ M[s][H_2] = M[s][H_1]. \]

    Now we can repeatedly apply \Cref{lemma:levy_elementary} (a) to get
    \[ M[G] = M[G_0][G_1][G_2] = M[s][H_0][G_1][G_2] = M[s][H_1][G_2] = M[s][H_2][G_2] \]
    and it happens that \(H_2 \cup G_2\) is of the form in \Cref{lemma:levy_elementary} (a), and thus generic over \(M[s]\).
\end{proof}




\section{Random Reals}

\newcommand*{\B}{{\mathcal{B}}}

Here we follow set-theoretic practice and associate ``reals'' with the Cantor set \(2^\omega\).
The definitions and results presented apply to any other standard space,
such as Baire space \(\omega^\omega\) and real line \(\mathbb{R}\), mutatis mutandis.

\begin{definition}
    Let \(s \in 2^{<\omega}\) be a binary string, denote \(O(s) = \set{f\in 2^\omega: f\supseteq s}\).
\end{definition}
\begin{proposition}
    \(\set{O(s): s\in 2^{<\omega}}\) is a base for the topology on \(2^\omega\).
\end{proposition}
\begin{definition}
    For each \(s\in 2^{<\omega}\), the Lebesgue measure of its corresponding \emph{basic open set} is defined as
    \[ m_L(O(s)) = \frac{1}{2^{\abs{s}}}. \]
\end{definition}

As \(2^{<\omega}\) is countable, for the rest of this chapter
we fix an enumeration \[ \Angle{\mathbf{s}_i : i\in\omega} \]
where the enumeration is nice enough in the sense that \(\abs{\mathbf{s}_i} \leq i\),
sequences appear after their proper initial segments, and the enumeration is computable.

We have been forcing with partial functions,
this time for a change we force with the \(\sigma\)-algebra of Borel sets.
Let \[ \B^* = \set{ X: X \text{ is a non-null Borel set}} \]
be ordered by inclusion.

\begin{definition}[Coding Borel sets]
    Each \(c:\omega\to\omega\) can encode a Borel set as follows
    \[ A_c = \begin{cases} \begin{aligned}
        &\bigcup \set{O(\mathbf{s}_i): c(i+1)=1} &\text{if } c(0) = 0, \\
        &2^\omega\setminus \bigcup\set{O(\mathbf{s}_i): c(i+1)=1} &\text{if } c(0) = 1, \\
        &\bigcap_{n\in\omega} \bigcup\set{O(\mathbf{s}_i) : c(2^n3^{i+1})=1} &\text{otherwise}.
    \end{aligned} \end{cases} \]
\end{definition}

Our coding system has the following properties,
\begin{enumerate}[i.]
    \item If \(c(0) = 0\), \(A_c\) is open;
    \item if \(c(0) = 1\), \(A_c\) is closed;
    \item if \(c(0) > 1\), \(A_c\) is \(G_\delta\).
    \item Each open, closed and \(G_\delta\) set is indexed by some code \(c\).
\end{enumerate}
So if \(c\) is the code for an open (resp. closed, \(G_\delta\)) set,
we call \(c\) an \emph{open (resp. closed, \(G_\delta\)) code}.

\begin{remark}
    In Solovay's original proof \autocite{solovay1970}, he encodes all Borel sets using a more powerful scheme.
    However, we know that every Borel set can be approximated by some \(G_\delta\) set,
    so a simpler coding scheme that indexes all \(G_\delta\) sets suffices to prove the results we need.
\end{remark}

\begin{proposition} Let \(M\) be a transitive model of \ZF, then
    \begin{enumerate}[(a)]
        \item The following notions are absolute for \(M\):
            \[ \begin{matrix}
                A_c &;& A_c = 0 &;& A_c\subseteq A_d &;& A_c\subseteq\paren{2^\omega\setminus A_d} &;& A_c\cap A_d.
            \end{matrix} \]
        \item The Lebesgue measure is absolute in the sense that for any code \(c\in M\),
            \[ m_L^M(A_c^M) = m_L(A_c).\]
    \end{enumerate}
\end{proposition}
\begin{proof}
    For part (a) note that our enumeration \(\Angle{\mathbf{s}_i}\) only mentions finitist objects so it is absolute,
    the \(\sigma\)-algebra operations are defined in terms of \(O(\mathbf{s}_i)\), and are therefore a composition of absolute notions.

    For (b) we case-split on \(c\),
    \begin{enumerate}
        \item If \(c\) is an open code, then \(A_c\) can be expressed as a disjoint union of basic open sets,
            in fact, we can effectively compute \(J\subseteq \omega\)
            such that \(A_c\) is written as a disjoint union \(A_c = \bigcup_{j\in J} \mathbf{s}_j\).
            Thus \(m_L(A_c)\) can be seen to be determined absolutely.
        \item If \(c\) is a closed code,
            we can compute a code \(d\) such that \(A_d = 2^\omega\setminus A_c\) which is open,
            then apply (1.) and (a).
        \item If \(c\) is a \(G_\delta\) code,
            we can effectively determine a sequence of open codes \(\Angle{d_n:n\in\omega}\)
            such that \(A_c = \bigcap_n A_{d_n}\).
            Once again we can reuse (1.) and determine \(m_L(A_c) = \inf_{n\in\omega} m_L(A_{d_n})\) absolutely.
            \hfill\qedhere
    \end{enumerate}
\end{proof}

Now we have an important characterization of \(\B^*\)-forcing.

\begin{theorem} \label{theorem:random_real}
    Suppose \(G\) is \(\B^*\)-generic,
    then there is a unique real \(x \in 2^\omega\) such that for any closed code \(c\in M\),
    \[ x\in A_c^{M[G]} \text{ iff } A_c^M \in G, \]
    and \(M[x] = M[G]\).
\end{theorem}
\begin{proof}
    Since every Borel set can be approximated by a closed set, for each \(n\in\omega\), the set
    \begin{equation} \label{eqn:borel_algebra_dense}
        \set{C\in \B^*: C\text{ is closed} \land \paren{\exists k\in2}\paren{\forall f\in C}\paren{f(n) = k} }\quad\text{ is dense.}
    \end{equation}
    Also, for any Lebesgue measurable set \(A\subseteq 2^\omega, A\in M\), the set
    \begin{equation} \label{eqn:borel_algebra_dense2}
        \set{C\in \B^*: C\text{ is closed} \land \paren{C\subseteq A\lor C\cap A = 0} }\quad\text{ is dense.}
    \end{equation}

    Work in \(M[G]\), there is a unique real \(x\in2^\omega\) specified by
    \[ \set{x} = \bigcap\set{A_c^{M[G]}: c\in M \text{ is a closed code and } A_c^M\in G} \]
    where the intersection is nonempty due to \(G\) being a filter and the result is a singleton due to \Cref{eqn:borel_algebra_dense}.

    For any closed code \(c\in M\).  If \(A_c^M\in G\) then \(x\in A_c^{M[G]}\) by definition of \(x\).
    Conversely suppose \(x\in A_c^{M[G]}\),
    to show that \(A_c^M\in G\),
    by \Cref{eqn:borel_algebra_dense2}
    we just have to show that \(A_c^M\) meets every closed set in \(G\).
    Let \(d\) be a closed code with \(A_d^M\in G\),
    we have \(x \in A_c^{M[G]} \cap A_d^{M[G]}\) by definition of \(x\),
    so by absoluteness, \(x \in A_c^M \cap A_d^M\) and \(A_c^M\) meets every closed set in \(G\) as desired.

    We have already shown that \(x\) is definable from \(M[G]\),
    to show \(M[G]\subseteq M[x]\) we define \(G\) in terms of \(x\) as
    \[ G = \set{p\in\B^* : \exists c\paren{c\in M\text{ is a closed code} \land x\in A_c^{M[x]} \land A_c^{M[x]} \subseteq p} } \]
    which shows that \(M[x] = M[G]\).
\end{proof}

\begin{definition}
    The unique real \(x\) that is added via \(\B^*\)-forcing is called a \emph{random real}.
    We say \(x\) is \emph{random over} \(M\) if \(x\) satisfies the conditions in \Cref{theorem:random_real}
    for some \(\B^*\)-generic \(G\).
\end{definition}

The following characterisation gives us insight into the etymology of the term ``random real''.
The Cantor set \(2^\omega\) behaves like a universal probability space in some ways, and
this characterisation draws a parallel between \(\B^*\)-forcing and avoiding measure zero (probability zero) sets,
hence random.
\begin{theorem} \label{theorem:bstar_forcing_avoid_null_set}
    A real \(x\in 2^\omega\) is random over \(M\) iff
    \(x\notin A_c\) for any \(c\in M\) which is a \(G_{\delta}\) code for a null set.
\end{theorem}
\begin{proof}
    Suppose \(x\) is random over \(M\) with \(G\) the witnessing \(\B^*\)-generic filter.
    Let \(c\in M\) be a \(G_\delta\) code for a null set, we want \(x\notin A_c\).
    Applying standard fact of measure theory, in \(M\) we have
    \[ D = \set{ C\in\B^*: C\text{ closed} \land C\subseteq \paren{2^\omega\setminus A_c} }\quad \text{ is dense.} \]
    By genericity, \(G\) meets \(D\) so let \(d\in M\) be a closed code such that \(A_d^M \in G\cap D\),
    then \(x\in A_d^{M[G]}\) by definition of \(x\),
    so by absoluteness, \(x\in A_d\) which means \(x\notin A_c\).

    Conversely suppose for all \(c\in M\) which is a \(G_\delta\) code for a null set, \(x\notin A_c\).
    We recover \(G\) from \(x\) in the manner of proof of \Cref{theorem:random_real}
    and show that \(G\) is \(\B^*\)-generic over \(M\).
    This means we just need to show that whenever \(D\in M\) is dense in \(\B^*\),
    there exists \(p\in D\) and a closed code \(c\in M\) satisfying \(x\in A_c\) and \(A_c^M \subseteq p\).

    Proceed with a dense \(D\in M\), let \(\mathcal{A}\in M\),
    \(\mathcal{A}\subseteq \set{C: C\text{ closed} \land \exists p\in D\paren{C\subseteq p}}\)
    be a maximal antichain.
    We know that the \(\B^*\) algebra is c.c.c. so \(\abs{\mathcal{A}} \leq \omega\).
    Let \(\Angle{ c_n:n \in\omega}\) be a sequence of closed codes such that
    \(\Angle{ A_{c_n} : n\in \omega }\) enumerates \(\mathcal{A}\).
    We have refined the information contained in the dense set \(D\) into something easier to work with,
    that is an antichain of closed sets enumerated by a countable sequence of closed codes.

    Now define \(c\) to be a \(G_\delta\) code (\(c(0) > 1\)) such that
    \[ c(2^n3^{i+1}) = 1 \text{ iff } O(\mathbf{s}_i) \cap A_{c_n} = 0. \]
    For each \(n\in\omega\), \(A_c\) includes all basic open sets that avoids \(A_{c_n}\),
    so \[A_c = \bigcap_{n\in\omega} \paren{2^\omega\setminus A_{c_n}} = 2^\omega\setminus\paren{\bigcup_{n\in\omega} A_{c_n}}.\]
    As \(\mathcal{A}\) is maximal, \(A_c\) has to be a null set.

    By hypothesis, \(x\notin A_c\), then \(x\in A_{c_n}\) for some closed code \(c_n\in M\) as desired.
\end{proof}


\section{Solovay's Theorem}

For this section, we assume \(\kappa\in M\) and \(M\models \kappa \text{ is an inaccessible cardinal}\).

\begin{definition}
    A set \(X\) is \(\mathbf{ORD}^\omega\)-definable, or \emph{definable from a countable sequence of ordinals} iff
    for some \(s:\omega\to\mathbf{ORD}\) and some formula \(\phi(x_1,x_2)\),
    \[ y\in X \text{ iff } \phi(s,y). \]
\end{definition}
\begin{remark}
    Formally, to avoid quantifying over all logical formulas in the metatheory, we use the reflection principle.
\end{remark}

The methods used in \autocite{myhillscott1971} allow us to prove this fact.
\begin{proposition}
    The collection of all hereditarily \(\mathbf{ORD}^\omega\)-definable sets
    %\(\mathbf{HCOD} = \set{X: \trcl(X) \subseteq \mathbf{COD}}\)
    is an inner model of \(\ZF\).
\end{proposition}

We first need to establish a key lemma which is built on the previously-established properties of the Lévy collapse.
\begin{lemma} \label{lemma:solovay_key}
    For each formula \(\phi(v)\), there is a \(\tilde{\phi}(v)\) such that
    for any \(s:\omega\to\mathbf{ORD}, s\in M[G]\),
    \[ M[G]\models \phi(s) \text{ iff } M[s] \models \tilde{\phi}(s). \]
\end{lemma}
\begin{proof}
    For any \(s\in \mathbf{ORD}^\omega \cap M[G]\), by \Cref{lemma:important}
    there is a \(\P = \Lv(\kappa)\)-generic filter \(H\) over \(M[s]\) such that \(M[G] = M[s][H]\).
    Since \(\P\) is almost homogeneous, applying \Cref{proposition:property_almost_homo} with ground model \(M[s]\) we have
    \[ M[s][H]\models \phi(s) \text{ iff } M[s] \models \left\ulcorner 1_\P \forces \phi(\check{s}) \right\urcorner. \]
    As forcing is definable in the ground model,
    we can take \(\tilde{\phi}\) to be the statement with one free variable that encapsulates the forcing assertion.
\end{proof}

\begin{theorem}
    Let \(G\) be \(\Lv(\kappa)\)-generic,
    then in \(M[G]\), each subset of reals definable from a countable sequence of ordinals is Lebesgue measurable.

    In particular, there is a \textbf{Solovay model} \(N\), with \(M \subseteq N \subseteq M[G]\),
    containing precisely all hereditarily \(\mathbf{ORD}^\omega\)-definable sets where every subset of reals is Lebesgue measurable.
\end{theorem}
\begin{proof}
    Work in \(M[G]\),
    for any \(s\in\mathbf{ORD}^\omega\),
    by \Cref{lemma:levy_elementary} (d) \(\kappa\) remains inaccessible in \(M[s]\),
    so \(\paren{\abs{\omega^\omega} < \kappa}^{M[s]}\) which gives us that
    \[ \omega^\omega \cap M[s] \text{ is countable}. \]
    Suppose \(A\subseteq 2^\omega\) is \(\mathbf{ORD}^\omega\)-definable,
    then for some \(s:\omega\to\mathbf{ORD}\) and \(\phi(v_1,v_2)\),
    \[ x\in A \text{ iff } \phi(s,x). \]
    Applying \Cref{lemma:solovay_key} we have another formula \(\tilde{\phi}(v_1,v_2)\) satisfying
    \[ x\in A \text{ iff } M[s][x] \models \tilde{\phi}(s,x). \]
    Consider random-real forcing over \(M[s]\), by \Cref{theorem:bstar_forcing_avoid_null_set} we have
    \[
        \set{x\in2^\omega: x\text{ is not random over } M[s]}
        = \bigcup \set{A_c: c\in M[s] \text{ is a } G_\delta \text{ code for a null set}}
    \]
    and the expression on the right is a countable union of null sets, so these sets are null.
    As Lebesgue measurable sets are characterised by being almost equal to Borel sets,
    we just need to find a Borel set \(X\) such that the symmetric difference \(A\triangle X\) only consists of non-random reals.

    For the \(\B^*\)-forcing argument over \(M[s]\),
    let \(\mathring{r}\) denote the canonical name for the random real that will be added.
    Now let \(Y \in M[s]\) be a maximal antichain consisting of closed sets
    that force either \(\tilde{\phi}(\check{s},\mathring{r})\) or its negation.
    If \(x\) is random over \(M[s]\),
    \(x\in A\) iff \(M[s][x]\models \tilde{\phi}(s,x)\)
    iff \(p\forces \tilde{\phi}(\check{s},\mathring{r})\) for some \(p\in H\) and \((\B^*)^{M[s]}\)-generic \(H\).
    As \(H\) is generic, it meets \(Y\) which by \Cref{theorem:random_real} gives us that
    \[ x\in A\text{ iff } x\in \bigcup\set{A_c: A_c^{M[s]}\in Y\land A_c^{M[s]} \forces \tilde{\phi}(\check{s},\mathring{r})}. \]
    By c.c.c. of \(\B\), \(Y\) is countable, so the expression on the right is in fact a Borel set,
    which witnesses \(X\).
\end{proof}

\section{Properties of Solovay Model}

\subsection{Dependent Choice}

The Axiom of Dependent Choice is a weaker form of Axiom of Choice that is sufficient to develop most of analysis.

\begin{axiom}[Dependent Choice]
    \[
        \forall X\forall R\paren{
            \begin{split}
                &X\ne0 \land R\subseteq X\times X \land \paren{\forall x\in X}\paren{\exists y\in X}\paren{\Angle{x,y}\in R} \\
                &\qquad \rightarrow \paren{\exists f\in X^\omega}\paren{\forall n\in\omega}\paren{\Angle{f(n),f(n+1)}\in R}
            \end{split}
        }
    \]
\end{axiom}
It turns out that Dependent Choice holds in the Solovay model, which shows that Dependent Choice is indeed a strictly weaker axiom.
\begin{lemma}
    Dependent Choice holds in \(N\).
\end{lemma}
\begin{proof}
    Let \(X,R\in N\) satisfy the hypotheses of Dependent Choice,
    applying \Axiom{Choice} in \(M[G]\),
    we can recursively define \(f:\omega\to X\) with \(\Angle{f(n),f(n+1)}\in R\) for all \(n\),
    which will be \(\mathbf{ORD}^\omega\)-definable, and hence be in \(N\).
\end{proof}

We can combine our results to make the following observation.
\begin{observation}
    The existence of non-measurable sets of real numbers is not provable from the Axiom of Dependent Choice.
\end{observation}

\subsection{Relative Consistency}

In the previous section, we started off from a base model \(M\) which contains a (strongly) inaccessible cardinal.
The consistency result obtained was
\[
    \begin{split}
        &\Con\paren{\ZFC + \exists \text{ strongly inaccessible cardinal}} \\
        &\qquad \rightarrow \Con\paren{\ZF + \lnot\Axiom{Choice} + \text{every subset of $\mathbb{R}$ is Lebesgue measurable}}.
    \end{split}
\]

In Gödel's constructible universe \(\mathbf{L}\), the Generalized Continuum Hypothesis holds,
so if \(\kappa\) is weakly inaccessible,
\[ \mathbf{L}\models \kappa\text{ is strongly inaccesible}. \]
This gives the consistency result that
\[
    \begin{split}
        &\Con\paren{\ZF + \exists\text{ weakly inaccessible cardinal}} \\
        &\qquad \rightarrow \Con\paren{\ZFC + \exists\text{ strongly inaccessible cardinal}}.
    \end{split}
\]

Combining the consistency results together,
this means we could have started from a ground model of only \ZF and a weakly inaccessible cardinal and prove the same result.

It is then natural to ask if the result can be proved without assuming the consistency of having an inaccessible cardinal,
Shelah in \autocite{shelah1984} showed that the assumption cannot be dropped.
